Canonical rings of $\bq$-divisors on projective space and Hirzebruch surfaces

### 2015-07-16 surfaces-write-up/surface-write-up.tex
edits to section 5

We define the $\beta_i$ above so that $\beta_i$ is the expansion of $f_i$ when viewed as a polynomial in $k[f_1,\ldots, f_{m+1}]$. Furthermore, the relations given in Equation ~\eqref{eqn:pm-relation-degree}
generate all relations, since they allow us to reduce any $\prod_{j =
1}^n f_j^{r_i}$ to a canonical form as given by Proposition ~\ref{prop:cone-generation}.

@@

\section{Canonical Rings on Hirzebruch surfaces}
Let $D=\sum_{i=1}^n \alpha_i D_i \in \mathbb{Q} \otimes \di(\hirz_
k).$ such
that $D_1, \ldots, D_4$ are distinct divisors with bi-degrees $(1,0)
, (1,0), (0,1)$, and $(0,1)$ respectively. Let $D_i = V(f_i)$ where
$f_i \in \sco(\bida_i, \bidb_i)$.  
Then, $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$. \todo{Talk a little about Hirzebruch surfaces before this paragraph; especially, describe what it's sections look like.}

@@

\begin{equation}
\label{eqn:define-T+(D)}
	\Tp(D) = \left\{ i \in \{1, \ldots, n\}:  \vphantom{\sum_{k = 1}^n} 
	\bida_i \sum_{k = 1}^m \alpha_k \bidb_k > \bidb_i \sum_{k = 1}^n \alpha_k \bida_k 
\right\},
\end{equation}

@@

Let $E$ be the set of $\epsilon_i$ and $\epsilon_{i,j}$ given 
above, satisfying the respective conditions.
The lattice points corresponding to elements of $E$ satisfies the conditions of Proposition ~\ref{prop:cone-generation}, so by applying that proposition, $R_D$ is generated in 
degrees less than
\[
	\sigma = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{\substack{
	i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}.
\]
\end{proof}

@@

To find such a $\beta$, we simply write $f_i$ as a 
polynomial in $f_1, \ldots f_4$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{j =
1}^n f_j^{r_j}$ to the canonical form (see Proposition ~\ref{prop:cone-generation}).  \todo{Peter: I think the canonical form given in the proposition isn't the one we actually want.}

Note that we can view the relations from Equation
~\ref{eqn:hirz-relations-psi} \todo{is this the right ref?} as an element of $S$ given
by $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$.  Let $S'\subseteq (\mathbb{R}_{\ge 0})^{n+1}$ be the
set of $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$ satisfying $c_i \ge -\alpha_i d$,
$\bida_j + \sum_{i = 1}^m \bida_i c_i = 0$, and $\bidb_j + \sum_{i=1}^n \bidb_i c_i
= 0$.  A careful computation deduces that $\gamma (d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots, c_n)$ is of the form 
\[
	E + \sum_{k\in \Te} r_k e_k + \sum_{\substack{h \in \Tp \\ k \in \Tm}} r_{h,k} e_{h,k}
\]

>>>>

We define the $\beta_i$ above so that $\beta_i$ is the expansion of $f_i$ when viewed as a polynomial in $k[f_1,\ldots, f_{m+1}]$. Furthermore, the relations given in Equation ~\eqref{eqn:pm-relation-degree}
generate all relations, since they allow us to reduce any $\prod_{j =
1}^n f_j^{r_i}$ to a canonical form, with $r_i = -\lfloor d \alpha_i\rfloor$ whenever $i  > m + 1$.

@@

\section{Canonical Rings on Hirzebruch surfaces}
Let $D=\sum_{i=1}^n \alpha_i D_i \in \mathbb{Q} \otimes \di(\hirz_
m).$ such
that $D_1, \ldots, D_4$ are distinct divisors with bi-degrees $(1,0)
, (1,0), (0,1)$, and $(0,1)$ respectively. Let $D_i = V(f_i)$ where
$f_i \in \sco(\bida_i, \bidb_i)$.  
Then, $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$. \todo{Talk a little about Hirzebruch surfaces before this paragraph; especially, describe what it's sections look like.}

@@

\begin{equation}
\label{eqn:define-T+(D)}
	\Tp(D) = \left\{ i \in \{1, \ldots, n\}:  \vphantom{\sum_{k = 1}^n} 
	\bida_i \sum_{k = 1}^n \alpha_k \bidb_k > \bidb_i \sum_{k = 1}^n \alpha_k \bida_k 
\right\},
\end{equation}

@@

The set of $e_i$ and $e_{i,j}$ given above satisfies the conditions of Proposition ~\ref{prop:cone-generation}, so by applying that proposition, $R_D$ is generated in 
degrees less than
\[
	\sigma = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i + \sum_{\substack{
	i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j- \bida_j \bidb_i)\ell_{i,j}.
\]
\end{proof}

@@

To find such a $\beta$, we simply write $f_i$ as a 
polynomial in $f_1, \ldots f_4$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{j =
1}^n f_j^{r_j}$ to the canonical form where $r_i = - \lfloor d \alpha_i \rfloor$ whenever $i > 4$.

Note that we can view a relation from Equation
~\ref{eqn:hirz-relations-psi} as an element of $S$ given
by $(d, c_1, \ldots, c_{i-1}, c_i + 1, c_{i+1}, \ldots , c_n)$.  Let $S'\subseteq (\mathbb{R}_{\ge 0})^{n+1}$ be the
set of $(d, r_1, \ldots, r_{i-1}, r_i + 1, r_{i+1}, \ldots , r_n)$ satisfying $r_i \ge -\alpha_i d$,
$\bida_i + \sum_{j = 1}^m \bida_j c_j = 0$, and $\bidb_i + \sum_{j=1}^n \bidb_j c_j
= 0$.  A careful computation deduces that $\gamma \in S'$ is of the form 
\[
	E + \sum_{k\in \Te} r_k e_k + \sum_{\substack{h \in \Tp \\ k \in \Tm}} r_{h,k} e_{h,k}
\]
